Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-03-03 src/tex/ms.tex
minor GW170818 tweaks

It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$, leads us to infer a reduced support for precession.
We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the time-dependent signal enhancement or dampening due to birefringence can mimic the amplitude modulation in a precession cycle.
Therefore, under these circumstances, similar morphologies can be obtained by setting $\chi_p > 0$ or $\kappa > 0$, as long as the distance and inclination can also be adjusted accordingly.

The fact that the birefringent analysis favors a high $\kappa$ for rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and of the phasing of the precession cycle.
The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR}.
We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.

>>>>

It is difficult to unequivocally identify a specific feature of the GW170818 data that leads to this posterior structure.
However, it appears to be related to this event's support for precession, in conjunction with its uncommonly definite measurement of the polarization and spin angles \cite{Varma:2021csh}.
The relevance of precession is evident from the posterior on $\chi_p$ (Fig.~\ref{fig:hist_GW170818}): allowing for $\kappa \neq 0$ leads to reduced support for precession.
We can understand this in reference to Fig.~\ref{fig:birefringence}: if only a short portion of the signal is observed, then the frequency-dependent signal enhancement or dampening due to birefringence can mimic the time-dependent amplitude modulation in a precession cycle.
Therefore, under these circumstances, similar morphologies can be obtained by setting $\chi_p > 0$ or $\kappa > 0$, as long as the distance and inclination can also be adjusted accordingly.

The fact that the birefringent analysis favors a high $\kappa$ rather than a high $\chi_p$ is a consequence of prior volume: many more configurations are available with long distances and high $\kappa$ than with short distances and small $|\kappa|$.
The preference for $\kappa > 0$ over $\kappa < 0$ (and, therefore, the lack of bimodality in the $\cos\iota$ and $\kappa$ posteriors), is likely related to both the definite measurement of left-handed polarizations and the specific phasing of the precession cycle.
The latter manifests as a precise constraint on the spin orientation and phase angles in the \ac{GR} analysis \cite{Varma:2021csh}.
The observed amplitude modulation (say, increasing vs decreasing towards the merger) likely determines the allowed sign of $\kappa$ for this event.
Indeed, the inferred polarization angle in the birefringent case is notably different from the \ac{GR} one.
We provide the full corner plot for all relevant parameters in Fig.~\ref{fig:corner_GW170818_appendix} in Appendix XX.
